---
title: "Checkout Weather Data"
output: html_notebook
---

## Import and Clean Up

```{r}
library(dplyr, warn.conflicts = FALSE)
rawData <- read.csv('weather_data.csv')
# subset to data we have
rawData <- rawData[,!is.na(rawData[1, ])] %>% 
    arrange(DATE)
# format date
rawData$DATE <- as.Date(rawData$DATE)
head(rawData)
```

## A Few Exploratory Graphs

What percent of days were rainy? (PRCP > 0)

```{r}
rainDays <- rawData %>% 
    filter(PRCP > 0) %>% 
    nrow()
rainDays/nrow(rawData)
```

How much does it rain on the average "rainy day"?

```{r}
rawData %>% 
    filter(PRCP > 0) %>% 
    summarise(avgPRCP = mean(PRCP)) %>% 
    .$avgPRCP
```

Percentage of rainy days by month?

```{r}
rawData$RAIN <- ifelse(rawData$PRCP > 0, 1, 0)
rawData$MONTH_NUM <- as.numeric(format(rawData$DATE, '%m'))
rawData$MONTH_CHAR <- month.name[rawData$MONTH_NUM]

library(ggplot2)
rawData %>% 
    group_by(MONTH_NUM) %>% 
    summarise(pct_rain = sum(RAIN)/n()) %>% 
    ggplot(aes(x = MONTH_NUM, y = pct_rain, fill = pct_rain)) +
    geom_col() +
    scale_fill_gradient(low = 'light blue', high = 'dark blue')
```

Month matters, bring me summer!!

More chance of rain if it's rained the previous day?

```{r}
rawData$PREV_RAIN <- numeric(length = nrow(rawData))
rawData$PREV_RAIN[2:nrow(rawData)] <- rawData$RAIN[1:nrow(rawData)-1]

rawData %>% 
    group_by(PREV_RAIN) %>% 
    summarise(pct_rain = sum(RAIN)/n()) %>% 
    ggplot(aes(x = PREV_RAIN, y = pct_rain, fill = pct_rain)) +
    geom_col() +
    scale_fill_gradient(low = 'light blue', high = 'dark blue')
```

Suspicion confirmed!

Should be expext an interaction variable? Does the percent of rainy days following a rainy change throughout the year?

```{r}
rawData %>% 
    group_by(MONTH_NUM, PREV_RAIN) %>% 
    summarise(pct_rain = sum(RAIN)/n()) %>% 
    ggplot(aes(x = MONTH_NUM, y = pct_rain, color = as.factor(PREV_RAIN))) +
    geom_point()
```

The effect looks additive. It doesn't look like there needs to be an interaction term.

## Model 1

Build a data set to check model results:

```{r}
viewOutput <- data.frame(
    MONTH_CHAR = rep(month.name[1:12], 2),
    PREV_RAIN = c(rep(0, 12), rep(1, 12))
)
```


Probability of Rain

```{r}
probRain <- glm(RAIN ~ MONTH_CHAR + PREV_RAIN, data = rawData, family = "binomial")
summary(probRain)
```

```{r}
viewOutput %>% 
    mutate(ChanceOfRain = predict(probRain, newdata = viewOutput, type = 'response'))
```

Classification Model: Rain or No Rain

```{r}
rawData$RAIN_FACTOR <- as.factor(ifelse(rawData$RAIN==0, "No", "Yes"))

library(caret, warn.conflicts = FALSE)
inTrain <- createDataPartition(rawData$RAIN_FACTOR, p = 0.7, list = FALSE)
training <- rawData[inTrain, ]
testing <- rawData[-inTrain, ]
isRain <- train(RAIN_FACTOR ~ MONTH_CHAR + PREV_RAIN, data = training, method = "glm")
predictions <- predict(isRain, newdata = testing)
confusionMatrix(predictions, testing$RAIN_FACTOR)
```

```{r}
viewOutput %>% 
    mutate(isRain = predict(isRain, newdata = viewOutput))
```

71% accuracy and a little wonky. Apparently it will never rain tomorrow if it didn't rain today!

## Model 2

Is yesterday's temperature correlated with rain today's rain?

```{r}
rawData$PREV_LOW <- numeric(length = nrow(rawData))
rawData$PREV_LOW[2:nrow(rawData)] <- rawData$TMIN[1:nrow(rawData)-1]

rawData %>% 
    mutate(PREV_LOW = cut(PREV_LOW, 10),
           PREV_RAIN = as.factor(PREV_RAIN)) %>% 
    group_by(MONTH_NUM, PREV_LOW, PREV_RAIN) %>% 
    summarise(pct_rain = sum(RAIN)/n()) %>% 
    ggplot(aes(x = PREV_LOW, y = pct_rain, color = PREV_RAIN)) +
    geom_point() +
    facet_wrap('MONTH_NUM', nrow = 4)
```

Probability of Rain

```{r}
probRain2 <- glm(RAIN ~ MONTH_CHAR * PREV_LOW + PREV_RAIN, data = rawData, family = "binomial")
```

Classification Model: Rain or No Rain

```{r}
inTrain <- createDataPartition(rawData$RAIN_FACTOR, p = 0.7, list = FALSE)
training <- rawData[inTrain, ]
testing <- rawData[-inTrain, ]
isRain2 <- train(RAIN_FACTOR ~ MONTH_CHAR * PREV_LOW + PREV_RAIN, data = training, method = "glm")
predictions2 <- predict(isRain2, newdata = testing)
confusionMatrix(predictions2, testing$RAIN_FACTOR)
```

Nope.. same accuracty.

## Save Models and Data

```{r}
saveRDS(rawData, file = "app/data.rds")
saveRDS(probRain, file = "app/probRain.rds")
saveRDS(isRain, file = "app/isRain.rds")
```

